---
title: "Analysis"
author: "Ziyue"
date: "2026-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## library and import the data

```{r cars}
library(tidyverse)
library(ggplot2)
library(lme4)
highest_cloze_all <- read_csv("/Users/zliu/Documents/C03_C11/SE/process_and_clean/highest_cloze_all.csv")
colSums(is.na(highest_cloze_all))
```

## Visulize the cloze
```{r}
highest_cloze_all$Itemtype[highest_cloze_all$Itemtype == "CS"] <- "code_switching"
highest_cloze_all$Itemtype[highest_cloze_all$Itemtype == "E"]  <- "English"

# plot
set.seed(123)  # reproducible jitter

p <- ggplot(highest_cloze_all, aes(x = Itemtype, y = cloze_probability)) +
  geom_violin(
    aes(fill = Itemtype),
    trim = FALSE,
    alpha = 0.3
  ) +
  geom_point(
    aes(color = Itemtype),
    position = position_jitter(width = 0.08),
    size = 2.2,
    alpha = 0.85
  ) +
  stat_summary(
    fun = mean,
    geom = "point",
    color = "red",
    size = 4,
    shape = 19
  ) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.5,
    fatten = 3,
    color = "red"
  ) +
  facet_wrap(~ participant) +          # seperate screen by participant
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  scale_fill_manual(values = c(
    "English"    = "#3B6FB6",
    "CodeSwitch" = "#D97A00"
  )) +
  scale_color_manual(values = c(
    "English"    = "#3B6FB6",
    "CodeSwitch" = "#D97A00"
  )) +
  labs(
    title = "Highest cloze probability per item",
    x     = "Language condition",
    y     = "Max cloze Probability"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  )

ggsave(
  filename = "plot.png",
  plot = p,
  width = 8,
  height = 6)
```

## LME model
### prepare the data and check fixed effects
```{r}
# transfer to factor and set the baseline
highest_cloze_all <- highest_cloze_all %>%
  ungroup() %>%
  mutate(
    participant = relevel(as.factor(participant), ref = "human"),
    Itemtype = relevel(as.factor(Itemtype), ref = "E"))

# check whether there are fixed effects
# for comparison, we need to set REML = FALSE because different fixed effect can not be compared with TRUE

# null model
m0 <- lmer(cloze_probability ~ 1 + (1 | InnerNumber), data = highest_cloze_all, REML = FALSE)

# participant
m1 <- lmer(cloze_probability ~ participant + (1 | InnerNumber), data = highest_cloze_all, REML = FALSE)
anova(m0, m1) # check whether participant has effect -> Yes, there is a participant effect

# + Itemtype
m2 <- lmer(cloze_probability ~ participant + Itemtype + (1 | InnerNumber), data = highest_cloze_all, REML = FALSE)
anova(m1, m2)

# interaction (participant * Itemtype)
m3 <- lmer(cloze_probability ~ participant * Itemtype + (1 | InnerNumber), data = highest_cloze_all, REML = FALSE)
anova(m2, m3)
```
The anova results show that there are significant participant and Itemtype effets, and the interaction effect of participant and Itemtype.

### final LME model
```{r}
# in the final model, we change REML = TRUE
model <- lmer(cloze_probability ~ participant * Itemtype + (1 | InnerNumber), data = highest_cloze_all, REML = TRUE)
summary(model)
```

Participant effect: The highest cloze probability of human under English condition is 0.84, all LLMs have higher cloze probability than human (Claude = 0.16, GPT = 0.12, Mistral = 0.13), indicating that participant has significant effects and LLMs achieve higher cloze probability than human under English condition.

Itemtype effect: Humans achieve lower cloze probability under CS condition, compared to English condition (-0.076). 

Interaction effect: Compared to humans, Claude and Mistral show significantly less sensitivity to CS items.

