---
title: "individual differences cloze study"
output: html_notebook
date: october 9 2025
---
We want to know
(a) whether those with higher BLP (more Eng dominant, less Sp proficient, less switching) completed more or fewer non-common completions (more means that the underlying factor of lower cloze is not being proficient in Spanish; less means that the underlying factor may be experience with the function of switching)
(b) whether those with higher BCSP (more switching) completed more non-common completions (these tend to be more proficient in Spanish, too), so we cannot distinguish switching from spanish prof, but can at least see what the direction of these effects are


To do:
1. merge the spanish completion file with the highest cloze file;  merge with BLP/BCSP data
2. clean, such that there is a column for Engl+transltion, and subjects/items removed
3. code such that if response = highest = 1
4. subset to  CS completions only
6. glmer with BLP(centered?) as factor 
7. glmer with BCSP(centered?) as factor 



```{r libraries}
library(tidyverse)
library(lme4)
library(readxl)
```

# 1 reading and merging

```{r reading files and merging}

highestcloze <- read_excel("Cloze_SpEng_n122_highestcloze.xlsx", 
    sheet = "Cloze_SpEng_n122_highestcloze", 
    range = "A1:G283")
View(highestcloze)


indcloze <- read_excel("CS_Spanish_completion_exp_n156_2025-09-12_coded.xlsx")
View(indcloze)

demo <- read_csv("~/Prominence_Spanish/Analysis/BSPBCSP/BLP_BSCP_Lextale_n122.csv")
View(demo)


cloze <- merge(indcloze, highestcloze, by=c("InnerNumber", "Item", "Itemtype"))

cloze <- merge(cloze, demo, by="ID")

```

# cleaning

```{r cleaning}



# remove dropped participants



#making a vector of the kept ID numbers

kept_omitted <- read_table("Subjects_Kept_Omitted_2025-09-12.txt")
#View(kept_omitted)
colnames(kept_omitted) <- c("ID", "kept")
kept <- kept_omitted %>% 
  filter(kept ==1)
keptv <- unname(unlist(kept[,1]))


#taking omitted out of cloze data
cloze <- cloze %>% 
  filter (ID %in% keptv)

length(unique(cloze$ID))


#removing deleted items
# taking out problematic items:


drop_items=c(22, 83, 92, 127, 129, 133, 147, 69 )
#22 espectaculo is not the same as show
#83 imagines is not the same as pictures
#92 carta has ambiguity not equal to cards
#127 holster may not be known by all
#129 esquiaba may not be known by all (?)
#133 canilla not known by most
#147 has syntactic ambiguity in spanish, but not in English
#69 does not have "her" as beneficient in the Spanish version

#taking omitted out of cloze data

cloze <- cloze %>% 
  filter (!(InnerNumber %in% drop_items))

length(unique(cloze$InnerNumber))


```




# 3. code such that if response = highest = 1
Note that cloze_clean_wtrans is the highest completion
cloze_clean is the edited completino by the participant
English_translation is the English translation if the completion was in Spanish

```{r coding}

cloze<-cloze %>% 
  mutate(match =   case_when(
        cloze_clean_wtrans==cloze_clean ~ 1,
        cloze_clean_wtrans==English_translation ~ 1,
        TRUE ~ 0)) %>% 
  mutate(include = as.numeric(include))

  
  


```


```{r checking which items, part are invalid}
cloze %>% 
  mutate(ID=as.character(ID)) %>% 
  group_by(ID) %>% 
   summarize(mean=mean(include)) %>% 
  filter(mean==0)


cloze %>% 
   mutate(InnerNumber=as.character(ID)) %>% 
  group_by(InnerNumber) %>% 
   summarize(mean=mean(include)) %>% 
  filter(mean==0)




```

```{r filter out invalid responses}

cloze<-cloze %>% 
   filter(include ==1)   #filter out invalid responses
  
length(unique(cloze$InnerNumber))

length(unique(cloze$ID))

```


# 6. glmer with BLP(centered) as factor 

```{r subsetting}
#subsetting to CS only

clozeCS<-cloze %>%
    filter(Itemtype=="CS")


clozeCS$BLP.c = clozeCS$BLP - mean(clozeCS$BLP)
clozeCS$BCSP.c = clozeCS$BCSP - mean(clozeCS$BCSP)
clozeCS$SPLexTaleSalig.c = clozeCS$SPLexTaleSalig - mean(clozeCS$SPLexTaleSalig)



```

glmer

```{r BLP}
summary(model_match_BLP<- glmer(match~ BLP.c+(1|ID)+(1|InnerNumber), data=clozeCS, family=binomial)) #singularity
model_match_BLP<- glmer(match~ BLP.c+(1|InnerNumber), data=clozeCS, family=binomial)

summary(model_match_BLP)  #numerically: more English dominant (less switching, less SP proficient ), more matches, fewer other (!)
```


```{r BCSPP}
# model_match_BLP<- glmer(match~ BLP+(1|ID)+(1|InnerNumber), data=clozeCS, family=binomial) #singularity
model_match_BCSP<- glmer(match~ BCSP.c+(1|InnerNumber), data=clozeCS, family=binomial)

summary(model_match_BCSP)


model_match_BCSP_Lextale<- glmer(match~ BCSP.c+SPLexTaleSalig.c+(1|InnerNumber), data=clozeCS, family=binomial)

summary(model_match_BCSP_Lextale)   # more switching, more SP proficient , more matches, fewer other (!)
```


glmer try random effect

```{r BLP}
model_match_BLP_re<- glmer(match~ 1+(1|ID), data=clozeCS, family=binomial)
```


```{r BLP ranef}
ranef(model_match_BLP_re)
#we saved this amnually to an excel file
```


```{r read in match}

match_intercept <- read_excel("match_intercept.xlsx")
View(match_intercept)

match_intercept<- merge (match_intercept, demo, by="ID")

cor.test(match_intercept$match_intercept, match_intercept$BLP)   #nothing
cor.test(match_intercept$match_intercept, match_intercept$BCSP) #nothing
cor.test(match_intercept$match_intercept, match_intercept$SPLexTaleSalig) #positive!!! so more SP proficient more match
cor.test(match_intercept$match_intercept, match_intercept$SelfProfES) #noting 

```


# 7. Sanity check: doing thsi with English 

```{r subsetting E}
#subsetting to E only

clozeE<-cloze %>%
    filter(Itemtype=="E")


clozeE$BLP.c = clozeE$BLP - mean(clozeE$BLP)
clozeE$BCSP.c = clozeE$BCSP - mean(clozeE$BCSP)
clozeE$SPLexTaleSalig.c = clozeE$SPLexTaleSalig - mean(clozeE$SPLexTaleSalig)



```

glmer try random effect

```{r BLP}
model_matchE_BLP_re<- glmer(match~ 1+(1|ID), data=clozeE, family=binomial)
```


```{r BLP ranef}
ranef(model_matchE_BLP_re)
#we saved this manually to an excel file "matchE_intercept.xlsx"
```

```{r read in match}

matchE_intercept <- read_excel("matchE_intercept.xlsx")
View(matchE_intercept)

matchE_intercept<- merge (matchE_intercept, demo, by="ID")

cor.test(matchE_intercept$matchE_intercept, matchE_intercept$BLP)   #nothing
cor.test(matchE_intercept$matchE_intercept, matchE_intercept$BCSP) #nothing
cor.test(matchE_intercept$matchE_intercept, matchE_intercept$SPLexTaleSalig) #positive!!! so more SP proficient more match
cor.test(matchE_intercept$matchE_intercept, matchE_intercept$SelfProfES) #noting 

```



# 8. looking at both CS and English 

```{r cloze contrast}
cloze$Itemtype = as.factor(cloze$Itemtype)
cloze$BLP.c = cloze$BLP - mean(cloze$BLP)
cloze$BCSP.c = cloze$BCSP - mean(cloze$BCSP)
cloze$SPLexTaleSalig.c = cloze$SPLexTaleSalig - mean(cloze$SPLexTaleSalig)

contrasts(cloze$Itemtype) <- cbind(CS_E=c(-.5, .5))  #CS is negative

```

glmer try random effect

```{r lmer all}



model_match_all_Lextale<- glmer(match~ Itemtype*SPLexTaleSalig.c+(1|ID)+(1|InnerNumber), data=cloze, family=binomial)

summary(model_match_all_Lextale)



model_match_all_woxLextale<- glmer(match~ Itemtype+SPLexTaleSalig.c+(1|ID)+(1|InnerNumber), data=cloze, family=binomial)

summary(model_match_all_woxLextale)

anova(model_match_all_Lextale,model_match_all_woxLextale )

```

