---
title: "Analysis"
author: "Ziyue"
date: "2026-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## library and import the data

```{r cars}
library(tidyverse)
library(ggplot2)
library(lme4)
highest_cloze_all <- read_csv("/Users/zliu/Documents/C03_C11/GE/process_and_clean/highest_cloze_all.csv")
```

## Visulize the cloze
```{r}
# plot
set.seed(123)  # reproducible jitter

ggplot(highest_cloze_all, aes(x = Itemtype, y = cloze_probability)) +
  geom_violin(
    aes(fill = Itemtype),
    trim = FALSE,
    alpha = 0.3
  ) +
  geom_point(
    aes(color = Itemtype),
    position = position_jitter(width = 0.08),
    size = 2.2,
    alpha = 0.85
  ) +
  stat_summary(
    fun = mean,
    geom = "point",
    color = "red",
    size = 4,
    shape = 19
  ) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.5,
    fatten = 3,
    color = "red"
  ) +
  facet_wrap(~ participant) +          # seperate screen by participant
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  scale_fill_manual(values = c(
    "English"    = "#3B6FB6",
    "CodeSwitch" = "#D97A00"
  )) +
  scale_color_manual(values = c(
    "English"    = "#3B6FB6",
    "CodeSwitch" = "#D97A00"
  )) +
  labs(
    title = "Highest cloze probability per item",
    x     = "Language condition",
    y     = "Max cloze Probability"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold")
  )

```

## LME model
### prepare the data and check fixed effects
```{r}
# transfer to factor and set the baseline
highest_cloze_all <- highest_cloze_all %>%
  ungroup() %>%
  mutate(
    participant = relevel(as.factor(participant), ref = "human"),
    Itemtype = relevel(as.factor(Itemtype), ref = "English"))

# check whether there are fixed effects
# for comparison, we need to set REML = FALSE because different fixed effect can not be compared with TRUE

# null model
m0 <- lmer(cloze_probability ~ 1 + (1 | inner_number), data = highest_cloze_all, REML = FALSE)

# participant
m1 <- lmer(cloze_probability ~ participant + (1 | inner_number), data = highest_cloze_all, REML = FALSE)
anova(m0, m1) # check whether participant has effect -> Yes, there is a participant effect

# + Itemtype
m2 <- lmer(cloze_probability ~ participant + Itemtype + (1 | inner_number), data = highest_cloze_all, REML = FALSE)
anova(m1, m2)

# interaction (participant * Itemtype)
m3 <- lmer(cloze_probability ~ participant * Itemtype + (1 | inner_number), data = highest_cloze_all, REML = FALSE)
anova(m2, m3)
```
The anova results show that there are significant participant and Itemtype effets, and the interaction effect of participant and Itemtype.

### final LME model
```{r}
# in the final model, we change REML = TRUE
model <- lmer(cloze_probability ~ participant * Itemtype + (1 | inner_number), data = highest_cloze_all, REML = TRUE)
summary(model)
```

Participant effect: The highest cloze probability of human under English condition is 0.74, all LLMs have higher cloze probability than human (Claude = 0.26, GPT = 0.22, Mistral = 0.24), indicating that participant has significant effects and LLMs achieve higher cloze probability than human under English condition.

Itemtype effect: Humans are not sensitive to Itemtype (-0.009740). 

Interaction effect: compare to human, Claude and Mistral are not affected by the Itemtype, while GPT achieves lower cloze probability under code-switching condition (GPT + English = 0.741 + 0.22 = 0.963, GPT + code-switching = 0.963 - 0.0097 - 0.0665 ≈ 0.887).

